Active Learning for SAT Solver Benchmarking

### 2022-10-11 paper/paper.tex
Revise Conclusion
Mainly language, text made a bit shorter

In this work, we have discussed possible solutions to the \emph{New-Solver Problem}:
Given a new solver, we want to find its ranking amidst its competitors.
Our approaches provide accurate ranking predictions while only needing a fraction of the runtime resources that a full evaluation on all benchmarking instances would need.
We can determine a new solver's PAR-2 ranking with about \SI{92}{\%} accuracy, while only needing \SI{10}{\%} of the full-evaluation time.
We use a runtime discretization technique as this enables us to transform the regression problem of directly predicting runtimes into the much simpler notion of classification.
We have shown that, albeit being more simple, the classification of discrete runtime labels produces good results.
We have evaluated several ranking algorithms, instance-selection approaches, and stopping criteria within our sequential active-learning process.
%A model-uncertainty-based selection approach in combination with runtime-label-prediction ranking and a ranking-convergence stopping criterion showed the consistently best results in most cases.
%JB: "consistently best" -- it's the best approach for one delta, but not sure regarding other deltas nor its sensitivity to hyperparameters
%TF: commented out
We also took a brief look at which instance families are the most prevalent in selection decisions.

@@

In future work, it may be of interest to compare further ranking algorithms, instance-selection approaches, and stopping criteria.
Furthermore, it is possible to formulate the runtime discretization as an optimization problem.
Given a dataset, optimization could select the discretization technique with the best discriminatory power rather than selecting one approach upfront by hand.

A major shortcoming is the lack of parallelizability.
Our current approach selects instances one at a time.
Running benchmarks on a computing cluster with $n$ cores benefits from having a batch of $n$ instances at a time.
This is, however, not trivial since, on the one hand, a higher $n$ leads to less \emph{active learning} (because of bigger batch sizes) and, on the other hand, it is not clear how to synchronize the model update and instance selection without a barrier lock, which wastes a lot of runtime resources.

On a more general note, it would make sense to generalize this evaluation framework for arbitrary $\mathcal{NP}$-complete problems as all discussed active-learning methods are agnostic to the underlying problem.
Solely the runtime-prediction model might need some tweaking.
Those problems share most of the relevant properties of SAT solving, i.e., there are established problem-instance features, a full benchmark run takes days, and creating new problem solvers traditionally requires expert knowledge to hand-select instances of interest.

>>>>

In this work, we have discussed possible solutions to the \emph{New-Solver Problem}:
Given a new solver, we want to find its ranking amidst its competitors.
Our approach provides accurate ranking predictions while needing significantly less runtime that a full evaluation on a given benchmark set.
In particular, we can determine a new solver's PAR-2 ranking with about \SI{92}{\%} accuracy, while only needing \SI{10}{\%} of the full-evaluation time.
We have found that our framework produces good results working with discrete runtime labels rather than regressing actual runtimes, albeit the former being more simple.
We have evaluated several ranking algorithms, instance-selection approaches, and stopping criteria within our sequential active-learning process.
%A model-uncertainty-based selection approach in combination with runtime-label-prediction ranking and a ranking-convergence stopping criterion showed the consistently best results in most cases.
%JB: "consistently best" -- it's the best approach for one delta, but not sure regarding other deltas nor its sensitivity to hyperparameters
%TF: commented out
We also took a brief look at which instance families are the most prevalent in selection decisions.

@@

Future work may compare further ranking algorithms, instance-selection approaches, and stopping criteria.
Furthermore, it is possible to formulate runtime discretization as an optimization problem, directly selecting the most discriminatory discretization technique rather than doing an empirical comparison.

A major shortcoming of our current approach is the lack of parallelization, selecting instances one at a time.
Benchmarking on a computing cluster with $n$ cores benefits from having batches of $n$ instances.
However, bigger batch sizes $n$ impede \emph{active learning}.
Also, it is unclear how to synchronize model update and instance selection without wasting too much runtime.

On a more general note, one can apply our evaluation framework to arbitrary $\mathcal{NP}$-complete problems as all discussed active-learning methods are problem-agnostic.
Solely the runtime-prediction model might need some tweaking.
%JB: other components of framework might need tweaking as well, or this there something special about prediction model?
Those problems share most of the relevant properties of SAT solving, i.e., there are established problem-instance features, a full benchmark run takes days, and solver development traditionally requires expert knowledge to hand-select instances.
